---
title: "Assignment 02 - Data Viz and Wrangling"
author: "Juehan Wang"
date: "10/4/2021"
output: 
    html_document:
      toc: yes 
      toc_float: yes
      keep_md : yes 
    github_document:
      html_preview: false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(data.table)
library(tidyverse)
library(tidytext)
library(tibble)
library(dplyr)
library(leaflet)
```

## Learning objectives

Conducting data wrangling and visualize the data with key questions in mind.

## I. Data Wrangling

```{r download-datasets-merge, cache=TRUE}
fn1 <- "chs_individual.csv"
if (!file.exists(fn1))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", destfile = fn1)

chs_individual<-read.csv(fn1)
chs_individual<-as_tibble(chs_individual)

fn2 <- "chs_regional.csv"
if (!file.exists(fn2))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", destfile = fn2)

chs_regional<-read.csv(fn2)
chs_regional<-as_tibble(chs_regional)

chs <- merge(
  x = chs_individual,
  y = chs_regional,
  by = "townname"
)
```

## Step 1

After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.

In the case of missing values, impute data using the average within the variables “male” and “hispanic.”

```{r imputation}
summary(chs)
#var <- c("agepft", "height", "weight", "bmi", "fev", "fvc", "mmef")
#for (i in 1:length(var)){
chs <- chs %>% 
  group_by(male,race) %>% 
  mutate( agepft = ifelse(is.na(agepft), 
                         mean(agepft, na.rm = TRUE), 
                         agepft)) %>%
  mutate( height = ifelse(is.na(height), 
                         mean(height, na.rm = TRUE), 
                         height)) %>%
  mutate( weight = ifelse(is.na(weight), 
                         mean(weight, na.rm = TRUE), 
                         weight)) %>%
  mutate( bmi = ifelse(is.na(bmi), 
                         mean(bmi, na.rm = TRUE), 
                         bmi)) %>%
  mutate( fev = ifelse(is.na(fev), 
                         mean(fev, na.rm = TRUE), 
                         fev)) %>%
  mutate( fvc = ifelse(is.na(fvc), 
                         mean(fvc, na.rm = TRUE), 
                         fvc)) %>%
  mutate( mmef = ifelse(is.na(mmef), 
                         mean(mmef, na.rm = TRUE), 
                         mmef)) %>%
  mutate( no_24hr = ifelse(is.na(no_24hr), 
                         mean(no_24hr, na.rm = TRUE), 
                         no_24hr)) %>%
  mutate( pm2_5_fr = ifelse(is.na(pm2_5_fr), 
                         mean(pm2_5_fr, na.rm = TRUE), 
                         pm2_5_fr))
#}
summary(chs)
```

The number of rows is correct in the merged data.

All missing values in numeric variables have been imputed using the average within the variables “male” and “hispanic.”

## Step 2

Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24).

To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r obesity-level}
chs$obesity_level <- as.factor(fifelse(chs$bmi < 14, "underweight",
        fifelse(chs$bmi < 22, "normal",
                fifelse(chs$bmi <= 24, "overweight", "obese"))))

# summary table
chs %>%
  group_by(obesity_level) %>%
  summarize(min_bmi = min(bmi),
            max_bmi = max(bmi),
            Count = n()) %>%
  knitr::kable()
```

Due to the table, the variable "obesity_level" is rightly coded.

## Step 3

Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove”. The variable should have four categories in total.

Note: smoke_gas_exposure: "smoke_gas" is both “Second Hand Smoke” and “Gas Stove”, "smoke" is only “Second Hand Smoke”, "gas" is only “Gas Stove” and "neither" is neither “Second Hand Smoke” nor “Gas Stove”.

```{r smoke-gas-exposure}
#chs[, smoke_gas_exposure := fifelse(smoke == 1, fifelse(gasstove == 1, "smoke_gas", "smoke", na=NA) , fifelse(gasstove == 1, "gas", "neither", na=NA), na=NA)]

chs$smoke_gas_exposure <- as.factor(fifelse(chs$smoke == 1, fifelse(chs$gasstove == 1, "smoke_gas", "smoke", na=NA) , fifelse(chs$gasstove == 1, "gas", "neither", na=NA), na=NA))

summary(chs$smoke_gas_exposure)
```

## Step 4

Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiatory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure”.

```{r summary-tables}
# summary by town
chs[-which(is.na(chs$asthma)!="FALSE"),] %>%
  group_by(townname,asthma) %>%
  summarise(mean_fev = mean(fev),
            sd_fev = sd(fev),
            n = n(),
            sd_asthma = sd(asthma, na.rm = TRUE)) %>%
  mutate(prop = prop.table(n)) %>%
  knitr::kable()

# summary by sex
chs[-which(is.na(chs$asthma)!="FALSE"),] %>%
  group_by(male,asthma) %>%
  summarise(mean_fev = mean(fev),
            sd_fev = sd(fev),
            n = n(),
            sd_asthma = sd(asthma, na.rm = TRUE)) %>%
  mutate(prop = prop.table(n)) %>%
  knitr::kable()

# summary by obesity level
chs[-which(is.na(chs$asthma)!="FALSE"),] %>%
  group_by(obesity_level,asthma) %>%
  summarise(mean_fev = mean(fev),
            sd_fev = sd(fev),
            n = n(),
            sd_asthma = sd(asthma, na.rm = TRUE)) %>%
  mutate(prop = prop.table(n)) %>%
  knitr::kable()

# summary by “smoke_gas_exposure”
chs[-c(which(is.na(chs$smoke_gas_exposure)!="FALSE"),which(is.na(chs$asthma)!="FALSE")),] %>%
  group_by(smoke_gas_exposure,asthma) %>%
  summarise(mean_fev = mean(fev),
            sd_fev = sd(fev),
            n = n(),
            sd_asthma = sd(asthma, na.rm = TRUE)) %>%
  mutate(prop = prop.table(n)) %>%
  knitr::kable()
```

## II. Looking at the Data (EDA)

The primary questions of interest are: 1. What is the association between BMI and FEV (forced expiatory volume)? 2. What is the association between smoke and gas exposure and FEV? 3. What is the association between PM2.5 exposure and FEV?

Follow the EDA checklist from week 3 and the previous assignment. Be sure to focus on the key variables. Visualization Create the following figures and interpret them. Be sure to include easily understandable axes, titles, and legends.

### Check the dimensions, headers, footers, variable names and variable types.

```{r data-info}
dim(chs)
head(chs)
tail(chs)
str(chs)
```

### Take a closer look at the key variables.

```{r key-variables}
table(chs$townname)
summary(chs$bmi)
summary(chs$fev)
table(chs$smoke_gas_exposure)
table(chs$"PM2.5 exposure ?????")
```

## Step 1

Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r scatterplots}
chs %>%
  ggplot(mapping = aes(x = bmi, y = fev, color = townname)) + 
  geom_jitter() + 
  facet_grid(~townname)
```

The scatter plot shows that Long Beach has the lowest FEV (forced expiatory volume) and Apline has the highest FEV (forced expiatory volume).

## Step 2

Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r histograms, warning=FALSE}
# by BMI category
chs %>%
  ggplot(aes(y = fev, fill = obesity_level)) + 
  geom_histogram(bins = 50)

# by smoke/gas exposure
chs[-which(is.na(chs$smoke_gas_exposure)!="FALSE"),] %>%
  ggplot(aes(y = fev, fill = smoke_gas_exposure)) + 
  geom_histogram(bins = 50)
```

From the histograms, people of high BMI have the highest FEV (forced expiatory volume) while people who are underweight has the lowest. Plus, people who are exposed to gas only have the highest FEV (forced expiatory volume) while people who are exposed to both smoke and gas are more likely to have the lowest FEV (forced expiatory volume).

## Step 3

Barchart of BMI by smoke/gas exposure.

```{r Barchart}
chs[-which(is.na(chs$smoke_gas_exposure)!="FALSE"),] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = smoke_gas_exposure, y = fev, fill = smoke_gas_exposure), position = "dodge", stat="identity")
```

From the barchart we can conclude that people who are exposed to gas only have the highest FEV (forced expiatory volume) while people who are exposed to both smoke and gas are more likely to have the lowest FEV (forced expiatory volume).

## Step 4

Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

```{r summary-graphs}
chs[-which(is.na(chs$smoke_gas_exposure)!="FALSE"),] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = obesity_level, y = fev, fill = smoke_gas_exposure), position = "dodge", stat="identity") + 
  scale_fill_brewer(palette="Blues")

```

Based on the summary bar plot of FEV by BMI and FEV by smoke/gas exposure category, people who are obese and exposed to gas have the highest FEV, while people who are underweight and exposed to both smoke and gas have the lowest FEV. We can also conclude that people who have normal BMI have higher FEV compared to people who have higher or lower BMI. Plus, people who are exposed to gas are more likely to have higher FEV compared to other groups.

## Step 5

A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r map}
temp.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=chs$pm25_mass)
if (knitr::is_html_output()) {
leaflet(chs) %>%
   addProviderTiles('CartoDB.Positron') %>%
   addCircles(
   lat = ~lat, lng=~lon,
   label = ~paste0(round(pm25_mass,2), ' pm25_mass'), color = ~ temp.pal(pm25_mass),
   opacity = 1, fillOpacity = 1, radius = 500
     ) %>%
   # And a pretty legend
   addLegend('bottomleft', pal=temp.pal, values=chs$pm25_mass,
           title='PM2.5 mass', opacity=1)
} else {
  message("Sorry! No HTML.")
}
```



## Step 6

Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r visualization-association}

```

