---
title: "Assignment04"
author: "Juehan Wang"
date: "11/15/2021"
output: 
    html_document:
      toc: yes 
      toc_float: yes
      keep_md : yes 
    github_document:
      html_preview: false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r package, include=FALSE}
library(parallel)
```

# HPC

## Problem 1: Make sure your code is nice

Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  ans <- rowSums(mat)
  ans
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  ans <- t(apply(mat, 1, cumsum))
  ans
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)
```

From the test of speed, both of the new codes run faster.

## Problem 2: Make things run faster with parallel computing

The following function allows simulating PI

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():

```{r}
cl <- makePSOCKcluster(4L)
clusterSetRNGStream(cl,1231)
clusterExport(cl,"sim_pi")
system.time({
  ans <- unlist(parLapply(cl, rep(4000,10000),sim_pi))
  print(mean(ans))
})
```

Compared to the previous code, the code we rewrite runs faster.

# SQL

Setup a temporary database by running the following chunk

```{r, include=FALSE}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

## Question 1

How many movies is there available in each rating category.

```{r}
dbGetQuery(con, "
SELECT rating AS Rating, COUNT (*) AS Count
FROM film
GROUP BY rating")
```

The result of the number of movies available in each rating category is shown in the table.

## Question 2

What is the average replacement cost and rental rate for each rating category.

```{r}
dbGetQuery(con, "
SELECT rating AS Rating, AVG(replacement_cost) AS Replacement_cost_avg,
AVG(rental_rate) AS Rental_rate_avg
FROM film
GROUP BY rating
LIMIT 10")
```

The result of the average replacement cost and rental rate for each rating category is shown in the table.

## Question 3

Use table film_category together with film to find the how many films there are with each category ID

```{r}
dbGetQuery(con, "SELECT category_id AS Category_ID,
  COUNT (*) as Count
FROM film_category
GROUP BY category_id
LIMIT 10")
```

The result of the number of films with each first 10 category ID is shown in the table.

## Question 4

Incorporate table category into the answer to the previous question to find the name of the most popular category.

```{r}
dbGetQuery(con, "
SELECT film_category.category_id AS Category_id, category.name AS Name, COUNT(*) AS Count
FROM film_category
  INNER JOIN film ON film_category.film_id = film.film_id
  INNER JOIN category ON film_category.category_id = category.category_id
GROUP BY category.category_ID
ORDER BY count DESC
LIMIT 5
")
```

```{r}
dbDisconnect(con)
```

The result of the number of films with each first 10 category name is shown in the table.

Therefore, the most popular category is Sports.
